---
title: "Confidence Intervals"
author: "Luciano Lorenti"
date: "11/26/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require('PEIP')
```

# Theory

## $\mu$ unknown, bu $\sigma$ know
To get an impression of the expectation $\mu$, it is sufficient to give an estimate. The appropriate estimator is the sample mean:

$\hat{\mu} = \bar{X} = \frac{1}{n} \sum\limits_{i=1}^n X_i$

By standardizing, we get a random variable:

$Z = \frac{\bar{X} - \mu}{\alpha/\sqrt{n}} \sim N(0,1)$

dependent on the parameter $\mu$ to be estimated, but with a standard normal distribution independent of the parameter $\mu$. Hence it is possible to find numbers $-z$ and $z$, independent of $\mu$, between which $Z$ lies with probability $1 − \alpha$, a measure of how confident we want to be.

We take $1 − \alpha = 0.95$, for example. So we have:

$\begin{align}
P(-z \leq Z \leq z) &= 1-\alpha \\
= 1 - P(Z \geq z) - P(Z \leq -z) \\
\text{ Como la gaussiana es simétrica} \\
P(Z \geq z) &=  P(Z \leq -z) \\
\text{ entonces } \\
P(-z \leq Z \leq z) &= 1 - 2P(Z \geq z) = 1- \alpha \\
                   &= 2P(Z \geq z) = \alpha \\
                   &= P(Z \geq z) = \frac{\alpha}{2} \\
                   &= P(Z \geq z) = \frac{\alpha}{2}  \\
                   &= 1-P(Z \leq z) = \frac{\alpha}{2}  \\
                   &= 1-\frac{\alpha}{2}  = P(Z \leq z) \\
                   &= 1-\frac{\alpha}{2}  = P(Z \leq z) \\
\end{align}$

¿For wich number of $z$ happens that $1-\frac{\alpha}{2}  = P(Z \leq z)$ ?

Given that $Z$ follows a normal distribution $\Phi(z) = P(Z \leq z)$ then $z = \Phi^{-1}(\Phi(z)) \rightarrow z=\Phi^{-1}\left(1-\frac{\alpha}{2}\right)$

Now suppouse that we know the value of $z$, let us call it  $z_\alpha$ then

$\begin{align}
P(-z_\alpha \leq Z \leq z_\alpha) &= 1 - \alpha \\
P\left(-z_\alpha \leq \frac{\hat{X} - \mu}{\sigma / \sqrt{n}} \leq z_\alpha \right) = 1-\alpha \\
P\left(-z_\alpha \frac{\sigma}{\sqrt{n}} \leq \hat{X} - \mu \leq \frac{\sigma}{\sqrt{n}} z_\alpha \right) = 1-\alpha \\
P\left(-\hat{X} - z_\alpha \frac{\sigma}{\sqrt{n}} \leq - \mu \leq -\hat{X} + \frac{\sigma}{\sqrt{n}} z_\alpha \right) = 1-\alpha \\
P\left(\hat{X} + z_\alpha \frac{\sigma}{\sqrt{n}} \geq  \mu \geq \hat{X} - \frac{\sigma}{\sqrt{n}} z_\alpha \right) = 1-\alpha \\
P\left(\hat{X} - \frac{\sigma}{\sqrt{n}} z_\alpha \leq \mu \leq \hat{X} + z_\alpha \frac{\sigma}{\sqrt{n}}  \right) = 1-\alpha \\
\end{align}$

The limits of the interval is finally

$\left[\hat{X} - \frac{\sigma}{\sqrt{n}} z_\alpha, \hat{X} + z_\alpha \frac{\sigma}{\sqrt{n}}  \right]$

## $\mu$ unknown and $\sigma$ unknown

¿What happens when the \sigma is unknown?

We know that if $\hat{X}$ is the sample mean. $Z={\frac {{\overline {X}}_{n}-\mu }{\sigma /{\sqrt {n}}}}$ follows a Normal distribution with mean  0 and variance 1. But if we don't  the standar deviation we can think in a related quotient.

$T = \frac{\hat{X} - \mu}{S_n / \sqrt{n}}$ where $S_n^2 = \frac{1}{n-1} \sum\limits_{i=1}^n (x_i - \hat{x})^2$ then Z follows a t-Student distribution with n-1 degrees of freedom. 

# Excercises 

1. Find the value of $t_{n-1,\alpha/2}$ needed to construct a two sided confidence interval of the given level with the given sample size:
    Let $F(t)$ be the cumulative distribution of the t-student distribution
    a. Level 90%, sample size 12.
    
        $\begin{align}
            F(t) &= 1-\frac{0.1}{2} \\ 
            F^{-1}\left(1-\frac{0.1}{2}\right) &= t
        \end{align}$
        ```{r}
        qt(1-(0.1/2),  11)
        ```
    b. Level 95%, sample size 7.
        ```{r}
        qt(1-(0.05/2),  6)
        ```
    c. Level 99%, sample size 2.
        ```{r}
        qt(1-(0.01/2),  1)
        ```
    d. Level 95%, sample size 29.
        ```{r}
        qt(1-(0.05/2),  28)
        ```
2. Find the value of $t_{n−1,\alpha}$ needed to construct an upper or lower confidence bound in each of the situations in Exercise 1.
    $\begin{align}
        P(Z \leq z) = 1 - \alpha
    \end{align}$
    a. Level 90%, sample size 12.
    
        $\begin{align}
            F(t) &= 1-\frac{0.1}{2} \\ 
            F^{-1}\left(1-0.1\right) &= t
        \end{align}$
        ```{r}
        qt(1-(0.1),  11)
        ```
    b. Level 95%, sample size 7.
        ```{r}
        qt(1-(0.05),  6)
        ```
    c. Level 99%, sample size 2.
        ```{r}
        qt(1-(0.01),  1)
        ```
    d. Level 95%, sample size 29.
        ```{r}
        qt(1-(0.05),  28)
        ```
3. Find the level of a two-sided confidence interval that is based on the given value of $t_{n−1,\alpha/2}$ and the given sample size.

\begin{align}
    P(-t \leq Z \leq t) &=  1 - P(Z \geq t) - P(Z \leq -t) \\
                    c &=  1 - 2P(Z \geq t) \\
                    c &= 1-2(1-P(Z \leq t)) \\
                    c &= 1-2+2P(Z \leq t) \\
                    c &= -1+2P(Z \leq t)
                    
\end{align}
        
    
a. t = 2.776, sample size 5.
    
    ```{r}
     -1+(2*pt(2.776, 4)) 
    ```
b. t = 2.718, sample size 12.
    ```{r}
     -1+(2*pt(2.718, 11)) 
    ```
c. t = 5.841, sample size 4.
    ```{r}
     -1+(2*pt(5.841, 3)) 
    ```
d. t = 1.325, sample size 21.
    ```{r}
     -1+(2*pt(1.325, 20)) 
    ```
e. t = 1.746, sample size 17.
    ```{r}
     -1+(2*pt(1.746, 16)) 
    ```
    
4. True or false: The Student’st distribution may be used to construct a confidence interval for the mean of any population, so long as the sample size is small.

False

5. The article “Wind-Uplift Capacity of Residential Wood Roof-Sheathing Panels Retrofitted with Insulating Foam Adhesive” (P. Datin, D. Prevatt,
and W. Pang, Journal of Architectural Engineering, 2011:144–154) presents a study of the failure pressures of roof panels. Following are the failure
pressures, in kPa, for five panels constructed with 6d smooth shank nails. These data are consistent with means and standard deviations presented in
the article.

3.32 2.53 3.45 2.38 3.01

Find a 95% confidence interval for the mean failure pressure for this type of roof panel.

Let $\hat{X}$ be the mean of the data and $\hat{\sigma}$ the standar deviation then:
Let $Z = \frac{ \hat{X} - \mu}{\frac{\hat{\sigma}}{\sqrt{n}}}$ then Z follows a t-student distribution with n-1 degrees of freedom

\begin{align}
    P(-t \leq Z \leq t) &= 0.95 = 1 - 0.05 \\
    P(-t \leq \frac{\hat{X} - \mu}{\frac{\hat{\sigma}}{\sqrt{n}}} \leq t) &= 1 - 0.05 \\
    P(-t \frac{\hat{\sigma}}{\sqrt{n}} \leq \hat{X} - \mu \leq t \frac{\hat{\sigma}}{\sqrt{n}}) &= 1 - 0.05 \\
    P(-\hat{X} -t \frac{\hat{\sigma}}{\sqrt{n}} \leq - \mu \leq -\hat{X} + t \frac{\hat{\sigma}}{\sqrt{n}}) &= 1-0.5 \\
    P(\hat{X} +t \frac{\hat{\sigma}}{\sqrt{n}} \geq  \mu \geq \hat{X} - t \frac{\hat{\sigma}}{\sqrt{n}}) &= 1-0.5 \\
    P(\hat{X} - t \frac{\hat{\sigma}}{\sqrt{n}} \leq \mu \leq \hat{X} +t \frac{\hat{\sigma}}{\sqrt{n}} ) &= 1-0.5 \\
\end{align}

First we compute the mean and the standard deviation
```{r}
data = c(3.32, 2.53, 3.45, 2.38, 3.01)
X = mean(data)
sigma = sd(data)
```
Which is the value of t 
\begin{align}
    P(-t \leq Z \leq t) &= 0.95 = 1 - 0.05 \\
      &= 1- 2P(Z \geq t) \\
      &= 1- 2(1 -P(Z \leq t)) \\
      &= -1 +2P(Z \leq t) \\
      &= -1 +2F(t) \\
      c &= -1 + 2F(t) \\
      \frac{c+1}{2} &= F(t) \\
      F^{-1}\left(\frac{c+1}{2}\right) = t
\end{align}

```{r}
n = length(data)
t = qt((0.95 +1)/2, n-1)
t
```

```{r}
c(X-t*sigma/(sqrt(n)), X+t*sigma/(sqrt(n)))
```

6. The following are summary statistics for a data set.
Would it be appropriate to use the Student’s t distribution to construct a confidence interval from these data? Explain.

* N = 10
* Mean = 8.905
* Median = 6.105
* StDev = 9.690
* Minimum = 0.512
* Maximum = 39.920
* Q1 = 1.967
* Q3 = 8.103

No. The mean and the median are very different

7. The article “An Automatic Visual System for Marble Tile Classification” (L. Carrino, W. Polini, and S. Turchetta, Journal of Engineering Manufacture,
2002:1095–1108) describes a measure for the shade of marble tile in which the amount of light reflected by the tile is measured on a scale of 0–255. A perfectly black tile would reflect no light and measure 0, and a perfectly white tile would measure 255. A sample of nine Mezza Perla tiles were measured,
with the following results:

204.999 206.149 202.102 207.048 203.496 206.343 203.496 206.676 205.831
Is it appropriate to use the Student’s t statistic
to construct a 95% confidence interval for the mean
shade of Mezza Perla tile? If so, construct the confidence interval. If not, explain why not.
```{r}
data = c(204.999, 206.149, 202.102, 207.048, 203.496, 206.343, 203.496, 206.676, 205.831)
```
```{r}
qqnorm(data)
qqline(data)
```
```{r}
boxplot(data)
```

The data seems to be relativily normal and without any outliers.  It is possible to use the Student statistic.

```{r}
n = length(data)
t = qt((0.95 +1)/2, n-1)
c(X-t*sigma/(sqrt(n)), X+t*sigma/(sqrt(n)))
```


8. A chemist made eight independent measurements of  the melting point of tungsten. She obtained a sample mean of 3410.14 degrees Celsius and a sample standard deviation of 1.018 degrees.
a. Use the Student’s t distribution to find a 95% confidence interval for the melting point of ungsten.
```{r}
n = 8
X = 3410.14
sigma = 1.018
t = qt((0.95 +1)/2, n-1)
c(X-t*sigma/(sqrt(n)), X+t*sigma/(sqrt(n)))
```
b. Use the Student’s t distribution to find a 98% confidence interval for the melting point of tungsten.
```{r}
t = qt((0.98 +1)/2, n-1)
c(X-t*sigma/(sqrt(n)), X+t*sigma/(sqrt(n)))
```
c. If the eight measurements had been 3409.76, 3409.80, 3412.66, 3409.79, 3409.76, 3409.77, 3409.80, 3409.78, would the confidence intervals above be valid? Explain.

```{r}
data = c(3409.76, 3409.80, 3412.66, 3409.79, 3409.76, 3409.77, 3409.80, 3409.78)
n = length(data)
X = mean(data)
sigma = sd(data)
t = qt((0.98 +1)/2, n-1)
c(X, sigma)
```