---
title: "ConfidenceIntervalProportion"
author: "Luciano Lorenti"
date: "26/1/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
1. In a simple random sample of 70 automobiles registered in a certain state, 28 of them were found to have emission levels that exceed a state standard.
a. What proportion of the automobiles in the sample had emission levels that exceed the standard?
```{r}
print(28/70)
```
b. Find a 95% confidence interval for the proportion of automobiles in the state whose emission levels exceed the standard.
```{r}
n = 70
n = n +4
suc = 28 +2 
p = (suc)/(n)
z = qnorm((1+0.95)/2)
print(paste(p-z*sqrt(p*(1-p)/n), " - ", p + z*sqrt(p*(1-p)/n)))
```
c. Find a 98% confidence interval for the proportion of automobiles whose emission levels exceed the standard.
```{r}
n = 70
n = n + 4
suc = 28 +2 
p = (suc)/(n)
z = qnorm(1.98/2)
print(paste(p-z*sqrt(p*(1-p)/n), " - ", p + z*sqrt(p*(1-p)/n)))
```
d. How many automobiles must be sampled to specify the proportion that exceed the standard to within±0.10 with 95% confidence?
$$
P(p-0.10 \leq p \leq p + 0.10) = 0.95
$$

$$
\begin{align*}
\hat{p} + z_{\alpha/2}\sqrt{\dfrac{\hat{p}(1-\hat{p})}{n}} &= p + 0.10 \\
z_{\alpha/2}\sqrt{\dfrac{\hat{p}(1-\hat{p})}{n}} &=  0.10 \\
z_{\alpha/2}\dfrac{\sqrt{\hat{p}(1-\hat{p})}}{0.1} &=  \sqrt{n} \\
z_{\alpha/2}^2\dfrac{\hat{p}(1-\hat{p})}{0.1^2} &=  n \\
\end{align*}
$$
```{r}
n = 70
n = n +4 
suc = 28 +2 
p = (suc)/(n)
z = qnorm(1.95/2)
print(ceiling( (z/0.1)^2 * p*(1-p)  -4 ))
```
Note that I used the $p$ using the $n+4$ and success + 2 trick. But when we compute the $n$ we have to substract the additional 4 added.
e. How many automobiles must be sampled to specify the proportion that exceed the standard to within±0.10 with 98% confidence?
```{r}
n = 70
n = n +4 
suc = 28 +2 
p = (suc)/(n)
z = qnorm(1.98/2)
print(ceiling( (z/0.1)^2 * p*(1-p)  -4 ))
```
f. Someone claims that less than half of the automobiles in the state exceed the standard. With whatlevel of confidence can this statement be made?
$$
\begin{align*}
P \left(  p < 0.5 \right) &=  \alpha \\
P \left( \dfrac{p - \hat{p}}{\sqrt{\frac{\hat{p}(1-\hat{p})}{n}}} < \dfrac{0.5- \hat{p}}{\sqrt{\frac{\hat{p}(1-\hat{p})}{n}}} \right) &=  \alpha \\
P \left( Z < \dfrac{0.5- \hat{p}}{\sqrt{\frac{\hat{p}(1-\hat{p})}{n}}} \right) &=  \alpha \\
\end{align*} \\
$$
```{r}
n = 70
n = n +4 
suc = 28 +2 
p = (suc)/(n)
z = (0.5-p)/(sqrt((p*(1-p))/(n)))
print(pnorm(z))
```

2. During a recent drought, a water utility in a certaintown sampled 100 residential water bills and foundthat 73 of the residences had reduced their water con-sumption over that of the previous year.
```{r}
n = 100 + 4
suc = 73 + 2
p = suc/n
```
a. Find a 95% confidence interval for the proportion of residences that reduced their water consumption.
```{r}
z = qnorm(1.95/2)
s = sqrt(p*(1-p)/n)
paste(p-z*(s), " - ", p+z*s)
```
b. Find a 99% confidence interval for the proportion of residences that reduced their water consumption.
```{r}
z = qnorm(1.99/2)
s = sqrt(p*(1-p)/n)
paste(p-z*(s), " - ", p+z*s)
```
c. Find the sample size needed for a 95% confidenceinterval to specify the proportion to within ± 0.05.
$$
\begin{align*}
P\left( p-0.05 < p < p+0.05 \right) &= 0.95  \\
P\left( \hat{p}- z_{\frac{\alpha}{2}}\sqrt{\dfrac{\hat{p}(1-\hat{p})}{n}} <p<\hat{p} + z_{\frac{\alpha}{2}}\sqrt{\dfrac{\hat{p}(1-\hat{p})}{n}} \right) &= 0.95  \\
\hat{p} + 0.05 &= \hat{p} + z_{\frac{\alpha}{2}}\sqrt{\dfrac{\hat{p}(1-\hat{p})}{n}} \\
 0.05 &= z_{\frac{\alpha}{2}}\sqrt{\dfrac{\hat{p}(1-\hat{p})}{n}} \\
 \sqrt{n} &= \dfrac{ z_{\frac{\alpha}{2}}}{0.05}\sqrt{\hat{p}(1-\hat{p}}) \\
 n &= \left( \dfrac{ z_{\frac{\alpha}{2}}}{0.05} \right)^2 \hat{p}(1-\hat{p})
\end{align*}
$$
```{r}
n = 100 + 4
suc = 73 + 2
p = suc/n
z = qnorm(1.95/2)
n = ((z/0.05)^2)*(p*(1-p))
print(ceiling(n-4))
```
d. Find the sample size needed for a 99% confidenceinterval to specify the proportion to within ± 0.05.
```{r}
n = 100 + 4
suc = 73 + 2
p = suc/n
z = qnorm(1.99/2)
n = ((z/0.05)^2)*(p*(1-p))
print(ceiling(n-4))
```
e. Someone claims that more than 70% of residences reduced their water consumption. With what level of confidence can this statement be made?

$$
\begin{align*}
P\left( x >0.7 \right) &= \alpha \\
1- P\left(  x< 0.7 \right) &= \alpha \\
1-P\left(  x< 0.7 \right) &= \alpha  \\
1-P\left( \dfrac{x-\hat{p}}{\sqrt{\dfrac{\hat{p}(1-\hat{p})}{n}}} < \dfrac{0.7 - \hat{p}}{\sqrt{\dfrac{\hat{p}(1-\hat{p})}{n}}} \right) &= \alpha  \\
1-P\left( Z < \dfrac{0.7 - \hat{p}}{\sqrt{\dfrac{\hat{p}(1-\hat{p})}{n}}} \right) &= \alpha \\
\end{align*}
$$

```{r}
n = 100 + 4
suc = 73 + 2
p = suc/n
z = (0.7 - p)*sqrt(p*(1-p)/n)
print(1-pnorm(z))
```



f. If 95% confidence intervals are computed for 200 towns, what is the probability that more than 192 of the confidence intervals cover the true proportions?
$X  \sim Bin(200, 0.95)$. 

$$ P(X  > 192) = 1-P(x<192)$$

 $X$ can be approximated with $N(np, np(1-p))$
 
```{r}
n = 200
p = 0.95
u = n*p
s = n*p*(1-p)
z = (192 - u)/(sqrt(s))
print(1-pnorm(z))
print(1-pbinom(192, n, p))
```


3. A soft-drink manufacturer purchases aluminum cansfrom an outside vendor. A random sample of 70 cansis selected from a large shipment, and each is testedfor strength by applying an increasing load to the sideof the can until it punctures. Of the 70 cans, 52 meetthe specification for puncture resistance.
a. Find a 95% confidence interval for the proportionof cans in the shipment that meet the specification.
```{r}
n = 70 + 4
suc = 52 + 2
p = suc/n
z = qnorm(1.95/2)
paste(p - z*sqrt((p*(1-p))/n), " - ", p + z*sqrt((p*(1-p))/n))
```
b. Find a 90% confidence interval for the proportionof cans in the shipment that meet the specification.
```{r}
n = 70 + 4
suc = 52 + 2
p = suc/n
z = qnorm(1.90/2)
paste(p - z*sqrt((p*(1-p))/n), " - ", p + z*sqrt((p*(1-p))/n))
```
c. Find the sample size needed for a 95% confidenceinterval to specify the proportion to within±0.05.
```{r}
n = 70 + 4
suc = 52 + 2
p = suc/n
z = qnorm(1.95/2)
n = (z/0.05)^2 * p*(1-p)
print(ceiling(n-4))
```
d. Find the sample size needed for a 90% confidenceinterval to specify the proportion to within±0.05.
```{r}
n = 70 + 4
suc = 52 + 2
p = suc/n
z = qnorm(1.90/2)
n = (z/0.05)^2 * p*(1-p)
print(ceiling(n-4))
```
e. If a 90% confidence interval is computed each dayfor 300 days, what is the probability that more than 280 of the confidence intervals cover the true proportions?
Let $X \sim Bin(300, 0.9)$, I want to know $P(X > 280) = 1-P(X \leq 280)$
If $X \sim Bin(n, p)$ then $X \approx N(np, np(1-p))$. Let's call $Y \sim N(np, np(1-p))$ and 
$Z \sim N(0, 1)$.

$$
P(X \leq 280) \approx P(Y \leq 280) = P\left(Z \leq \dfrac{280 - np}{\sqrt{np(1-p)}}\right)
$$


```{r}
n = 300
p = 0.9
u = n*p
s = n*p*(1-p)
z = (280 - u)/(sqrt(s))
print(1-pnorm(280, mean = n*p, sd=sqrt(s)))
print(1-pnorm(z))
print(1-pbinom(280, 300, p))
```

4. The article “HIV-positive Smokers Considering Quit-ting: Differences by Race/Ethnicity” (E. Lloyd-Richardson, C. Stanton, et al.,Am J Health Behav,2008:3–15) surveyed 444 HIV-positive smokers. Ofthese, 170 reported that they had used a nicotine patch.Consider this to be a simple random sample.

```{r}
n = 444 + 4
suc = 170 + 2
p = suc /n
print(p)
```
a. Find a 95% confidence interval for the proportionof HIV-positive smokers who have used a nicotinepatch.
```{r}
n = 444 + 4
suc = 170 + 2
p = suc /n
z = qnorm(1.95/2)
dt = z*sqrt(p*(1-p)/n)
paste(p - dt, " - ",p + dt)
```
b. Find a 99% confidence interval for the proportionof HIV-positive smokers who have used a nicotinepatch.
```{r}
n = 444 + 4
suc = 170 + 2
p = suc /n
z = qnorm(1.99/2)
dt = z*sqrt(p*(1-p)/n)
paste(p - dt, " - ",p + dt)
```
c. Someone claims that the proportion is less than 0.40. With what level of confidence can this state-ment be made?
$$
\begin{align*}  
P\left( p < 0.4 \right) &= \alpha \\
P\left( \dfrac{p - \hat{p}}{\sqrt{\dfrac{\hat{p}(1-\hat{p})}{n}}} < \dfrac{0.4 - \hat{p}}{\sqrt{\dfrac{\hat{p}(1-\hat{p})}{n}}} \right) &= \alpha \\
P\left( Z < \dfrac{0.4 - \hat{p}}{\sqrt{\dfrac{\hat{p}(1-\hat{p})}{n}}} \right) &= \alpha \\
\end{align*}
$$

```{r}
n = 444 + 4
suc = 170 + 2
p = suc /n
z = (0.4 -p )/sqrt(p*(1-p)/n)
print(pnorm(z))
```
d. Find the sample size needed for a 95% confi-dence interval to specify the proportion to within±0.03.
$$
\begin{align*}
P\left(\hat{p} - 0.03 < p < \hat{p}+0.03 \right) &= 0.95 \\
P\left(\hat{p} - z_{\frac{\alpha}{2}}\sqrt{\dfrac{\hat{p}(1-\hat{p})}{n}} < p < \hat{p} + z_{\frac{\alpha}{2}}\sqrt{\dfrac{\hat{p}(1-\hat{p})}{n}} \right) &= 0.95 \\
\hat{p} + 0.03 = \hat{p} + z_{\frac{\alpha}{2}}\sqrt{\dfrac{\hat{p}(1-\hat{p})}{n}} \\
 0.03 =  z_{\frac{\alpha}{2}}\sqrt{\dfrac{\hat{p}(1-\hat{p})}{n}} \\
 \sqrt{n} =  \dfrac{z_{\frac{\alpha}{2}}}{0.03}\sqrt{\hat{p}(1-\hat{p})}  \\
 n =  \left( \dfrac{z_{\frac{\alpha}{2}}}{0.03}\right)^2 \hat{p}(1-\hat{p})
\end{align*}
$$
```{r}
n = 444 + 4
suc = 170 + 2
p = suc /n
z = qnorm(1.95/2)
n = ((z/0.03)^2)*(p*(1-p))
print(ceiling(n - 4))
```
e. Find the sample size needed for a 99% confidence interval to specify the proportion to within±0.03.
```{r}
n = 444 + 4
suc = 170 + 2
p = suc /n
z = qnorm(1.99/2)
n = ((z/0.03)^2)*(p*(1-p))
print(ceiling(n - 4))
```

5. The article “The Functional Outcomes of Total Knee Arthroplasty” (R. Kane, K. Saleh, et al., Journal of Bone and Joint Surgery, 2005:1719–1724) reports that out of 10,501 surgeries, 859 resulted in complications within six months of surgery.
a. Find a 95% confidence interval for the proportion of surgeries that result in complications within six months.
```{r}
n = 10501 + 4
suc = 859 + 2
p = suc/n
z = qnorm(1.95/2)
dt = z*sqrt(p*(1-p)/n)
paste(p-dt, " - ", p+dt)
```
b. Find a 99% confidence interval for the proportion of surgeries that result in complications within six months.
```{r}
n = 10501 + 4
suc = 859 + 2
p = suc/n
z = qnorm(1.99/2)
dt = z*sqrt(p*(1-p)/n)
paste(p-dt, " - ", p+dt)
```
c. A surgeon claims that the rate of complications is less than 8.5%. With what level of confidence can this claim be made?

$$
\begin{align*}
P\left(x < 0.085 \right) &= \alpha \\
P\left( \dfrac{x -\hat{p}}{\sqrt{\dfrac{\hat{p}(1-\hat{p})}{n}}} < \dfrac{0.085 -\hat{p}}{\sqrt{\dfrac{\hat{p}(1-\hat{p})}{n}}} \right) &= \alpha \\
P\left(Z < \dfrac{0.085 -\hat{p}}{\sqrt{\dfrac{\hat{p}(1-\hat{p})}{n}}} \right) &= \alpha \\
\end{align*}
$$
```{r}
n = 10501 + 4
suc = 859 + 2
p = suc/n
z = (0.085 - p)/(sqrt(p*(1-p)/n))
print(pnorm(z))
```

9. A random sample of 400 electronic components manufactured by a certain process are tested, and 30 are found to be defective.
a. Let p represent the proportion of components manufactured by this process that are defective. Find a 95% confidence interval for p.
```{r}
n = 400 + 4
suc = 30 + 2
p = suc/n
z = qnorm(1.95/2)
dt = sqrt(p*(1-p)/n)
paste(p - z*dt, " - ", p + z*dt)
```
b. How many components must be sampled so that the 95% confidence interval will specify the proportion defective to within ±0.02?
$$
\begin{align*}
P\left( \hat{p}-0.02 < p < \hat{p}+0.02 \right) &= 0.95 \\
P\left( \hat{p} - z_{\frac{\alpha}{2}}\sqrt{\dfrac{\hat{p}(1-\hat{p})}{n}} < p < \hat{p} + z_{\frac{\alpha}{2}}\sqrt{\dfrac{\hat{p}(1-\hat{p})}{n}} \right) &= 0.95 \\
0.02 &= z_{\frac{\alpha}{2}}\sqrt{\dfrac{\hat{p}(1-\hat{p})}{n}} \\
\sqrt{n} &= \dfrac{z_{\frac{\alpha}{2}}}{0.02}\sqrt{\hat{p}(1-\hat{p})} \\
n &= \left( \dfrac{z_{\frac{\alpha}{2}}}{0.02}\right)^2 \hat{p}(1-\hat{p}) \\
\end{align*}
$$

```{r}
n = 400 + 4
suc = 30 + 2
p = suc/n
z = qnorm(1.95/2)
n = ((z/0.02)^2)*p*(1-p)
paste(ceiling(n -4))
```
```{r}
n = 400 + 4
suc = 30 + 2
p = suc/n
z = qnorm(1.95/2)
dt = sqrt(p*(1-p)/n)
paste(p - z*dt, " - ", p + z*dt)
```
c. (Hard) The company ships the components in lots of 200. Lots containing more than 20 defective components may be returned. Find a 95% confidence interval for the proportion of lots that will be returned.
$ X \sim Bin(200, p)$  then $p1 = P(X>20) = 1- P(X \leq 20)$

```{r}
n = 400 + 4
suc = 30 + 2
p = suc/n
z = qnorm(1.95/2)
dt = sqrt(p*(1-p)/n)
plb = p - z*dt
pub = p + z*dt

n = 200
lb1 = 1 - pbinom(20, n, plb)
ub1 = 1 - pbinom(20, n, pub)
paste(lb1, " - ", ub1)

lb1 = 1 - pnorm(20, mean=n*plb, sd=sqrt(n*plb*(1-plb)))
ub1 = 1 - pnorm(20, mean=n*pub, sd=sqrt(n*pub*(1-pub)))
paste(lb1, " - ", ub1)

```

