---
title: "Theorem Limit Central"
author: "Luciano Lorenti"
date: "1/17/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
1. Bottles filled by a certain machine are supposed to contain 12 oz of liquid. In fact the fill volume is random with mean 12.01 oz and standard deviation 0.2 oz. 

a. What is the probability that the mean volume of a random sample of 144 bottles is less than 12 oz?
```{r}
u = 12
x = 12.01
s = 0.2
n= 144
z = (x - u)/(s/sqrt(n))
print(z)
print(paste("The probability is ", pnorm(-z)))

```
b. If the population mean fill volume is increased to 12.03 oz, what is the probability that the mean volume of a sample of size 144 will be less than 12 oz?
```{r}
u = 12
x = 12.03
s = 0.2
n= 144
z = (x - u)/(s/sqrt(n))
print(z)
print(paste("The probability is ", pnorm(-z)))
```

2. A 500-page book contains 250 sheets of paper. The thickness of the paper used to manufacture the book has mean 0.08 mm and standard deviation 0.01 mm.
a. What is the probability that a randomly chosen book is more than 20.2 mm thick (not including the covers)?
```{r}
u = 250*0.08
x = 20.2
s = 0.01
n = 250

z = (x - u)/(s*sqrt(n))

print(paste("The probability is ", pnorm(z, lower.tail=FALSE)))
```
b. What is the 10th percentile of book thicknesses?
$$P(z < S_1) = 0.1 $$
We have the unknown distribution  of the thicknes of the page with mean $\mu=0.08$ and $\sigma=0.01$, and 250 random variables of that distribution:
The sum $S_n = X_1 + X_2 + ... + X_n$
Then  $\lim\limits_{n \rightarrow \infty} P\left( \dfrac{S_n - nu}{\sigma\sqrt{n}} \leq z \right) = \Phi(z)$

So $z= \dfrac{S_n - nu}{\sigma\sqrt{n}} $ follows a normal distribution. We want to know for  wich value of that random variable:
$$
\begin{align*}
P(z < S_1) &= 0.1  \\
S_1 &= \Phi^{-1}(0.1)  \\
\end{align*}
$$

```{r}
S1 = qnorm(0.1)
print(S1)
```
So
$$
\begin{align*}
\lim\limits_{n \rightarrow \infty} P\left( \dfrac{S_n - n\mu}{\sigma\sqrt{n}} \leq -1.281552 \right) = \Phi(-1.281552) &= 0.1 \\
\lim\limits_{n \rightarrow \infty} P\left( S_n - n\mu \leq -1.281552 \sigma \sqrt{n} \right) &= 0.1 \\
\lim\limits_{n \rightarrow \infty} P\left( S_n - n\mu \leq -1.281552 \sigma \sqrt{n} + n \mu \right) &= 0.1 \\
\end{align*}
$$

```{r}
S1 = qnorm(0.1)
u = 0.08
s = 0.01
n = 250
print(n*u + sqrt(n)*s*S1)
```

c. Someone wants to know the probability that a randomly chosen page is more than 0.1 mm thick. Is enough information given to compute this probability? If so, compute the probability. If not, explain why not

No


3. A commuter encounters four traffic lights each day onher way to work. Let $X$ represent the number of these that are red lights. The probability mass function of $X$ is as follows. 

$$ P(X=0) = 0.1  \hspace{5pt} | \hspace{5pt}
P(X=1) = 0.3   \hspace{5pt} | \hspace{5pt}
P(X=2) = 0.3   \hspace{5pt} |\hspace{5pt}
P(X=3) = 0.2  \hspace{5pt} | \hspace{5pt} 
P(X=4) = 0.1  \hspace{5pt} 
$$

What is the probability that in a period of 100 days,the average number of red lights encountered is morethan 2 per day?

Let first compute the mean and the variance of the distribution of the number of traffic lights.

$$
\begin{align*}
\mu &= \sum\limits_{i=0}^4 i P(X=i)  \\
 &= 1*0.3 + 2*0.3 + 3*0.2 + 4*0.1
\end{align*}
$$
```{r}
mu = 1*0.3 + 2*0.3 + 3*0.2 + 4*0.1
print(mu)
```

Let's compute the variance $Var[X] = E[X^2]  - E[X]^2$

$$
\begin{align*}
E[X^2] &= \sum\limits_{i=0}^4 i^2 P(X=i) \\
&= 1*0.3 + 4*0.3 + 9*0.2 + 16*0.1\\
\end{align*}
$$

```{r}
mu = 1*0.3 + 2*0.3 + 3*0.2 + 4*0.1
muu =  1*0.3 + 4*0.3 + 9*0.2 + 16*0.1
var = muu - mu^2
print(var)
```



```{r}
n = 100
mu = 1*0.3 + 2*0.3 + 3*0.2 + 4*0.1
muu =  1*0.3 + 4*0.3 + 9*0.2 + 16*0.1
var = muu - mu^2
print(pnorm(2, mean=mu, sd=sqrt(var/n),lower.tail=FALSE))
```

Other way
$$
\begin{align*}
P\left(\hat{X} > 2 \right) &= P \left( Z > \dfrac{\hat{X} - \mu}{\sigma/\sqrt{n}} \right) \\
                     &= P \left( Z > \dfrac{2 - \mu}{\sigma/\sqrt{n}} \right) \\
\end{align*}
$$

```{r}
X = 2
mu = 1*0.3 + 2*0.3 + 3*0.2 + 4*0.1
muu =  1*0.3 + 4*0.3 + 9*0.2 + 16*0.1
var = muu - mu^2
z = (X-mu)/(sqrt(var/n))
print(pnorm(z,lower.tail=FALSE))
```

4. Among all the income-tax forms filed in a certain year,the mean tax paid was \$2000 and the standard devia-tion was \$500. In addition, for 10% of the forms, the tax paid was greater than \$3000. A random sample of 625 tax forms is drawn.

a. What is the probability that the average tax paidon the sample forms is greater than \$1980?
Every payment is a random varialbe, so $\hat{X} = \dfrac{X_1 + X_2 + ... + X_{625}}{n}$ follows approximately a normal distribution with $\mu_\hat{X} = \mu$ and variance $ \sigma^2_{\hat{X}} = \dfrac{\sigma^2}{n}$
```{r}
u = 2000
s = 500
n = 625
print(pnorm(1980, mean =u, s=s / sqrt(n), lower.tail=FALSE))

```
b. What is the probability that more than 60 of thesampled forms have a tax of greater than \$3000?

Lets consider the random variable $Y =$ The tax paid was greater than 3000. The probability is $p=0.1$
We want to see the probability of 
$$ P(Y_1 + Y_2 +...+Y_{625} > 60) $$
```{r}
p = 0.1
print(pnorm(60/625, mean=p, sd=p*(1-p)/sqrt(n), lower.tail=FALSE))
```

5. Bags checked for a certain airline flight have a meanweight of 15 kg with a standard deviation of 5 kg. A random sample of 60 bags is drawn

a. What is the probability that the sample meanweight is less than 14 kg?
```{r}
s = 5
n = 60
t = (14 - 15)/(5/sqrt(n))
print(pnorm(t, lower.tail=TRUE))
```
b. Find the 70th percentile of the sample mean weights.
$$
\begin{align*}
P\left( \dfrac{\hat{X} - \mu}{\sigma / \sqrt{n}} < z\right) &= \Phi(z) = 0.7 \\
z = \Phi^{-1}(0.7)
\end{align*}
$$
```{r}
z = qnorm(0.7)
print(z)
```

$$
\begin{align*}
P\left( \dfrac{\hat{X} - \mu}{\sigma / \sqrt{n}} < 0.5244005 \right) &=  0.7 \\
P\left(\hat{X}  < 0.5244005 \dfrac{\sigma}{\sqrt{n}} + \mu \right) &=  0.7 \\
\end{align*}
$$
```{r}
z = qnorm(0.7)
s = 5
n = 60
mu = 15
print(z*(s/sqrt(n)) + mu)

```

c. How many bags must be sampled so that the prob-ability is 0.01 that the sample mean weight is less than 14 kg?
$$
\begin{align*}
P\left( \dfrac{\hat{X} - \mu}{\sigma / \sqrt{n}} < z \right) &=  \Phi(z) = 0.01 \\
z_{0.01} &= \Phi^{-1}(0.01) \\
P\left( \dfrac{14 - 15}{5 / \sqrt{n}} < z_{0.01} \right) &=   0.01 \\
P\left( \dfrac{-1\sqrt{n}}{5 } < z_{0.01} \right) &=   0.01 \\
P\left( \sqrt{n} > - z_{0.01} 5 \right) &=   0.01 \\
P\left( n > (- z_{0.01} 5)^2 \right) &=   0.01 \\
\end{align*}
$$
```{r}
z = qnorm(0.01)
s = 5
print((z*s)^2)
```
Necesito 136 samples.

6. The amount of warpage in a type of wafer used in themanufacture of integrated circuits has mean 1.3 mm and standard deviation 0.1 mm. A random sample of 200 wafers is drawn.

a. What is the probability that the sample meanwarpage exceeds 1.305 mm?
```{r}
mu = 1.3
X = 1.305
s = 0.1
n = 200
z = (X-mu)/(s/sqrt(n))
print(pnorm(z, lower.tail=FALSE))
```
b. Find the 25th percentile of the sample mean.
```{r}
mu = 1.3
s = 0.1
n = 200
z = qnorm(0.25)
print(z*(s/sqrt(n)) + mu)
```
c. How many wafers must be sampled so thatthe probability is 0.05 that the sample meanexceeds 1.305?
$$
\begin{align*}
P\left(\dfrac{\hat{X} - \mu}{\sigma/\sqrt{n}} < z\right) &= \Phi(z) = 0.05 \\
z_{0.05} &= \Phi^{-1}(0.05) \\
P\left( \hat{X} - \mu < z_{0.05}\dfrac{\sigma}{\sqrt{n}} \right) &= 0.05 \\
P\left( \sqrt{n} < z_{0.05}\dfrac{\sigma}{\hat{X} - \mu} \right) &= 0.05 \\
P\left( n < \left(z_{0.05}\dfrac{\sigma}{\hat{X} - \mu}\right)^2 \right) &= 0.05 \\
\end{align*}
$$
```{r}
mu = 1.3
s = 0.1
n = 200
z = qnorm(0.05)
print(((z  * s)/(1.305 - mu))^2)
```

7. The time spent by a customer at a checkout counterhas mean 4 minutes and standard deviation 2 minutes.

a. What is the probability that the total time taken bya random sample of 50 customers is less than 180minutes?
$$
\begin{align*}
P\left( \dfrac{\hat{X} - \mu}{\sigma / \sqrt{n}} < z \right) = \Phi(z) \\
P\left( \dfrac{\hat{X} - \mu}{\sigma / \sqrt{n}}\dfrac{n}{n} < z \right) = \Phi(z) \\
P\left( \dfrac{S_X - \mu n}{\sigma \sqrt{n}} < z \right) = \Phi(z) \\
\end{align*}
$$
```{r}
mu = 4
Sx = 180
n = 50
s = 2
t  = (Sx-(mu*n))/(s*sqrt(n))
print(pnorm(t, lower.tail=TRUE))
```
b. Find the 30th percentile of the total time taken by  50 customers
```{r}
z = qnorm(0.3)
print(z*s*sqrt(n)  + mu*n)
```
8. Drums labeled 30 L are filled with a solution froma large vat. The amount of solution put into eachdrum is random with mean 30.01 L and standarddeviation 0.1 L.

a. What is the probability that the total amountof solution contained in 50 drums is morethan 1500 L?
```{r}
n = 50
X = 1500
s = 0.1
t = (1500 - 30.01*n)/(s*sqrt(n))
print(t)
print(pnorm(t, lower.tail=FALSE))
```
b. If the total amount of solution in the vat is 2401 L,what is the probability that 80 drums can be filledwithout running out?
$$
\begin{align*}
P\left( \sum\limits_{i=1}^{n} X_i \leq V \right) = P\left( \dfrac{\sum\limits_{i=1}^{n} X_i - \mu n}{\sigma \sqrt{n}}  \leq \dfrac{V - \mu n}{\sigma \sqrt{n}} \right)  = \Phi \left(\dfrac{V - \mu n}{\sigma \sqrt{n}} \right) \\
\end{align*}
$$
```{r}
n = 80
V = 2401
u = 30.01
s = 0.1
t = (V - u*n)/(s*sqrt(n))
print(t)
print(pnorm(t, lower.tail=TRUE))
```
c. How much solution should the vat contain so that the probability is 0.9 that 80 drums can be filledwithout running out?
$$
\begin{align*}
P \left( \sum\limits_{i=1}^n X_i \leq V \right) = 0.9 \\
P \left( \dfrac{\sum\limits_{i=1}^n X_i - n \mu}{\sigma \sqrt{n}} \leq \dfrac{V - n \mu}{\sigma \sqrt{n}} \right) &= 0.9 = \Phi\left(\dfrac{V - n \mu}{\sigma \sqrt{n}}\right) \\
\Phi^{-1}(0.9) &= \dfrac{V - n \mu}{\sigma \sqrt{n}} \\
\Phi^{-1}(0.9) &= \dfrac{V - n \mu}{\sigma \sqrt{n}} \\
\Phi^{-1}(0.9)\sigma \sqrt{n} + n \mu &= V
\end{align*}
$$

```{r}
n = 80
u = 30.01
s = 0.1
print(qnorm(0.9)*s*sqrt(n) + n*u)
```



9. The temperature of a solution will be estimated by taking n independent readings and averaging them.Each reading is unbiased, with a standard deviationof σ=0.5◦C. How many readings must be taken sothat the probability is 0.90 that the average is within±0.1◦C of the actual temperature?
  
$$
\begin{align*}
P \left( \left\vert \hat{X}_n  \right\vert \leq 0.1 \right) &= 0.9 \\
P \left( \left\vert \dfrac{\hat{X}_n}{\sigma/\sqrt{n}} \right\vert \leq \dfrac{0.1}{\sigma/\sqrt{n}} \right) &= 0.9 \\
P \left( \left\vert Z \right\vert \leq \dfrac{0.1}{\sigma/\sqrt{n}} \right) &= 0.9 \\
P \left( -\dfrac{0.1}{\sigma/\sqrt{n}} \leq  Z  \leq \dfrac{0.1}{\sigma/\sqrt{n}} \right) &= 0.9 \\
P \left( -\dfrac{0.1}{\sigma/\sqrt{n}} \leq  Z  \leq \dfrac{0.1}{\sigma/\sqrt{n}} \right) &= 0.9 \\
P\left(Z  \leq \dfrac{0.1}{\sigma/\sqrt{n}} \right) - P \left( Z \leq -\dfrac{0.1}{\sigma/\sqrt{n}} \right) &= 0.9 \\
\text{Ver la explicación mas abajo de porque } & P(Z \leq -v) = 1-P(Z \leq v) \\
P\left(Z  \leq \dfrac{0.1}{\sigma/\sqrt{n}} \right) - \left( 1- P \left( Z \leq
\dfrac{0.1}{\sigma/\sqrt{n}} \right) \right) &= 0.9 \\
2 P\left(Z  \leq \dfrac{0.1}{\sigma/\sqrt{n}} \right) - 1 &= 0.9 \\
P\left(Z  \leq \dfrac{0.1}{\sigma/\sqrt{n}} \right) &= \frac{1.9}{2} = \Phi \left( \dfrac{0.1}{\sigma/\sqrt{n}} \right) \\
\Phi^{-1}\left(  \frac{1.9}{2}  \right)  &= \dfrac{0.1 \sqrt{n}}{\sigma} \\
\Phi^{-1}\left(  \frac{1.9}{2}  \right) \frac{\sigma}{0.1}  &= \sqrt{n} \\
\left( \Phi^{-1}\left(  \frac{1.9}{2}  \right) \frac{\sigma}{0.1} \right)^2 &= n\\
\end{align*}
$$


```{r}
s = 0.5
v = 0.1
t = qnorm(1.9/2)
print(t)
print(ceiling(((t*s)/v)^2))
```

### $P(Z \leq -v) = 1-P(Z \leq v)$
Clearly the red area is equal to 1 - the blue area
```{r echo=FALSE}
mean=0; sd=1
lb=-555.1; ub=-1.1

x <- seq(-4,4,length=500)*sd + mean
hx <- dnorm(x,mean,sd)

plot(x, hx, type="n", ylab="",
  main="Normal Distribution", axes=FALSE)

i <- x >= lb & x <= ub
lines(x, hx)
polygon(c(lb,x[i],ub), c(0,hx[i],0), col="red")

area <- pnorm(ub, mean, sd) - pnorm(lb, mean, sd)
axis(1, at=seq(40, 160, 20), pos=0) 

lb1=1.1; ub1=5555.5
i <- x >= lb1 & x <= ub1
polygon(c(lb1,x[i],ub1), c(0,hx[i],0), col="blue")

area <- pnorm(ub1, mean, sd) - pnorm(lb1, mean, sd)
abline(v =lb1)
abline(v =ub)
mtext("v", at=lb1)
mtext("-v", at=ub)
```