---
title: "TestVariance"
author: "Luciano Lorenti"
date: "1/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. A random sample of size 11 from a normal distribution has variance $s^2 = 96$. 

Test $H_0 : \sigma^2 ≤ 50$ versus $H_1 : \sigma^2 > 50$.

We know  that $$(n-1) \dfrac{s^2}{\sigma^2} \sim \chi^2_{n-1}$$ If we asume $H_0$ true $\sigma^2 = 50$

```{r}
n = 11
s1 = 96
s2 = 50
chi = ((n-1)*s1)/s2
print(pchisq(chi, n-1, lower.tail=FALSE))

```


2. A random sample of size 29 from a normal distribution has variance $s^2 = 24$. 
Test $H_0 : \sigma^2 \geq 30$ versus $H_1 : \sigma^2 < 30$.

```{r}
n = 29
s1 = 24
s2 = 30
chi = ((n-1)*s1)/s2
print(pchisq(chi, n-1, lower.tail=TRUE))
```


3.Scores on an IQ test are normally distributed. A sample of 25 IQ scores had variance $s^2 = 64$. The developer of the test claims that the population variance is $\sigma^2 = 225$. Do these data provide sufficient evidence to contradict this claim?

$$H_0 : \sigma^2 = 255 \text{ versus } H_1 : \sigma^2 \neq 255$$

```{r}
n = 25
s1 = 64
s2 = 225
chi = ((n-1)*s1)/s2
p = pchisq(chi, n-1, lower.tail=TRUE) 
print(2*p)
```

4. A machine that fills beverage cans is supposed to put 12 ounces of beverage in each can. The variance of the amount in each can is 0.01. The machine is
moved to a new location. To determine whether the variance has changed, 10 cans are filled. Following are the amounts in the 10 cans. Assume them to be a
random sample from a normal population.

12.18  11.77   12.09  12.03  11.87 11.96  12.03  12.36  12.28  11.85

Perform a hypothesis test to determine whether the variance differs from 0.01. What do you conclude?
$$ H_0: \sigma^2 = 0.01 \text{ versus } H_1: \sigma^2 \neq 0.01 $$

```{r}
data = c(12.18, 11.77, 12.09, 12.03, 11.87, 11.96, 12.03, 12.36, 12.28, 11.85)
s1 = sd(data)^2
s2 = 0.01
n = length(data)
chi = (n-1)*(s1/s2)
plower = pchisq(chi, n-1, lower.tail=TRUE)
pupper = pchisq(chi, n-1, lower.tail=FALSE)
p = min(plower, pupper)
print(p*2)
```

5. A sample of 25 one-year-old girls had a mean weight of 24.1 pounds with a standard deviation of 4.3 pounds. Assume that the population of weights is
normally distributed. A pediatrician claims that the standard deviation of the weights of one-year-old girls is less than 5 pounds. Do the data provide convincing evidence that the pediatrician’s claim is true? (Based on data from the National Health Statistics Reports.)

```{r}
n = 25
s1 = 4.3^2
s2 = 5^2
chi = (n-1)*(s1/s2)
p = pchisq(chi, n-1, lower.tail=TRUE)
print(p)
```

6. The 2008 General Social Survey asked a large number of people how much time they spent watching TV each day. The mean number of hours was 2.98 with
a standard deviation of 2.66. Assume that in a sample of 40 teenagers, the sample standard deviation of daily TV time is 1.9 hours, and that the population of TV watching times is normally distributed.
Can you conclude that the population standard deviation of TV watching times for teenagers is less than 2.66?

```{r}
n = 40
s1 = 1.9^2
s2 = 2.66^2
chi = (n-1)*(s1/s2)
p = pchisq(chi, n-1, lower.tail=TRUE)
print(p)
```



7. Scores on the math SAT are normally distributed. A sample of 20 SAT scores had standard deviations = 87. Someone says that the scoring system for the SAT is designed so that the population standard deviation will be σ = 100. Do these data provide sufficient evidence to contradict this claim?

```{r}
n = 20
s1 = 87^2
s2 = 100^2
chi = (n-1)*(s1/s2)
plower = pchisq(chi, n-1, lower.tail=TRUE)
pupper = pchisq(chi, n-1, lower.tail=FALSE)
p = min(plower, pupper)
print(2*p)
```

8. One of the ways in which doctors try to determine how long a single dose of pain reliever will provide relief is to measure the drug’s half-life, which is the length of time it takes for one-half of the dose to be eliminated from the body. A report of the Naional Institutes of Health states that the standard
deviation of the half-life of the pain reliever oxycodone is σ = 1.43 hours. Assume that a sample of 25 patients is given the drug, and the sample standard  deviation of the half-lives was s = 1.5 hours. As sume the population is normally distributed. Can you conclude that the true standard deviation is greater than the value reported by the National Institutes of Health?

```{r}
n = 25
s1 = 1.5^2
s2 = 1.43^2
chi = (n-1)*(s1/s2)
p = pchisq(chi, n-1, lower.tail=FALSE)
print(p)
```

9. Find the upper 5% point of F7,20.
```{r}
print(qf(0.95, 7, 20), lower.tail=FALSE)
```
10.Find the upper 1% point ofF2,5.
```{r}
print(qf(0.99, 2, 5), lower.tail=FALSE)
```

11. An F test with five degrees of freedom in the numerator and seven degrees of freedom in the denominator produced a test statistic whose value was 7.46.
a. What is the P-value if the test is one-tailed?
```{r}
print(pf(7.46, 4, 6, lower.tail=FALSE))
```
b. What is the P-value if the test is two-tailed
```{r}
print(2*pf(7.46, 4, 6, lower.tail=FALSE))
```

12. A broth used to manufacture a pharmaceutical product has its sugar content, in mg/mL, measured several times on each of three successive days.

Day 1: 5.0 4.8 5.1 5.1 4.8 5.1 4.8 4.8 5.0 5.2 4.9 4.9 5.0
Day 2: 5.8 4.7 4.7 4.9 5.1 4.9 5.4 5.3 5.3 4.8 5.7 5.1 5.7
Day 3: 6.3 4.7 5.1 5.9 5.1 5.9 4.7 6.0 5.3 4.9 5.7 5.3 5.6

a. Can you conclude that the variability of the process is greater on the second day than on the first day?
```{r}
day_1 = c(5.0, 4.8, 5.1, 5.1, 4.8, 5.1, 4.8, 4.8, 5.0, 5.2, 4.9, 4.9, 5.0)
day_2 = c(5.8, 4.7, 4.7, 4.9, 5.1, 4.9, 5.4, 5.3, 5.3, 4.8, 5.7, 5.1, 5.7)

n1 = length(day_1)
s1 = var(day_1)

n2 = length(day_2)
s2 = var(day_2)

print(2*pt(s1/s2, n1-1, n2-1,lower.tail=FALSE))

```
b. Can you conclude that the variability of the process is greater on the third day than on the second day?

```{r}
day_1 = c(5.0, 4.8, 5.1, 5.1, 4.8, 5.1, 4.8, 4.8, 5.0, 5.2, 4.9, 4.9, 5.0)
day_2 = c(5.8, 4.7, 4.7, 4.9, 5.1, 4.9, 5.4, 5.3, 5.3, 4.8, 5.7, 5.1, 5.7)
day_3 = c(6.3, 4.7, 5.1, 5.9, 5.1, 5.9, 4.7, 6.0, 5.3, 4.9, 5.7, 5.3, 5.6)
```

```{r}
n1 = length(day_1)
s1 = var(day_1)

n2 = length(day_2)
s2 = var(day_2)

n3 = length(day_3)
s3 = var(day_3)
```

13.Refer to Exercise 11 in Section 5.6. Can you con-clude that the variance of the sodium content differsbetween the two brands?

Brand A:  34.36, 31.26,  37.36, 28.52, 33.14, 32.74, 34.34, 34.33, 30.95
Brand B:  41.08, 38.22, 39.59, 38.82, 36.24, 37.73,  35.03, 39.22, 34.13,34.33, 34.98, 29.64, 40.60

```{r}
brand_a = c(34.36, 31.26,  37.36, 28.52, 33.14, 32.74, 34.34, 34.33, 30.95)
brand_b = c(41.08, 38.22, 39.59, 38.82, 36.24, 37.73,  35.03, 39.22, 34.13,34.33, 34.98, 29.64, 40.60)
va = var(brand_a)
vb = var(brand_b)
na = length(brand_a)
nb = length(brand_b)
print(vb/va)
print(pf(vb/va, na-1, nb-1,  lower.tail=FALSE))
```