---
title: "T-Test"
author: "Luciano Lorenti"
date: "11/26/2019"
output: html_document
---
    
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
1. Each of the following hypothetical data sets represents some repeated weighings of a standard weight that isknown to have a mass of 100 g. Assume that the readings are a random sample from a population that follows the normal curve. we

a. 100.02, 99.98, 100.03
$$H_0 = 100 \hspace{10pt} H_1 \neq 100$$
```{r}
data = c(100.02, 99.98, 100.03)
s = sd(data)
N = length(data)
mu = mean(data)
t = (mu - 100)/(s/sqrt(N))
print(t)
```
Since we are using the the sample standard deviation we know that the statistic 
$$ T = \dfrac{\hat{X}_n  - \mu_0}{s/\sqrt{n}} \sim t_{n-1} \text{ under } H_0 $$ 
So our statistic follow the t-distribution of n-1 degrees of freedom. The p-value is the area of the curve at the right of $t$ and to the left of $t$. We have to use the $pt$ function:
```{r}
p_value = pt(-t, df=2, lower.tail = TRUE) + pt(t, df=2, lower.tail = FALSE)
print(p_value)
```
Since the p-value is a large value we cannot discard the null hypothesis.
b. 100.01
With one sample I can't do anything. I cant compute the std


2.  A geologist is making repeated measurements (in grams) on the mass of a rock. It is not known whether the measurements are a random sample from an approximately normal population. Below are three sets of replicate measurements, listed in the order they were made. For each set of readings, state whether
the assumptions necessary for the validity of the t test appear to be met. If the assumptions are not met, explain why
a. 213.03 212.95 213.04 213.00 212.99 213.01 221.03 213.05
```{r  fig.height = 4, fig.width = 9}
data = c(213.03, 212.95, 213.04, 213.00, 212.99, 213.01, 221.03, 213.05)
par(mfrow=c(1, 3))
boxplot(data)
hist(data, breaks = 15)
qqnorm(data); qqline(data)
```
    
This data doest not appear to follow a normal dsitribution

b. 213.05 213.00 212.94 213.09 212.98 213.02 213.06 212.99

```{r  fig.height = 4, fig.width = 9}
data = c(213.05, 213.00, 212.94, 213.09, 212.98, 213.02, 213.06, 212.99)
par(mfcol=c(1, 3))
boxplot(data)
hist(data, breaks = 15)
qqnorm(data); qqline(data)
```

c. 212.92 212.95 212.97 213.00 213.01  213.04 213.05 213.06

```{r  fig.height = 4, fig.width = 9}
data = c(212.92, 212.95, 212.97, 213.00, 213.01,  213.04, 213.05, 213.06)
par(mfrow=c(1, 3))
boxplot(data)
hist(data, breaks = 15)
qqnorm(data); qqline(data)
```


3. A new centrifugal pump is being considered for an application involving the pumping of ammonia. The specification is that the flow rate be more than 5 gallons per minute (gpm). In an initial study, eight runs were made. The average flow rate was 6.5 gpm and the standard deviation was 1.9 gpm. If the mean flow rate is found to meet the specification, the pump will be put into service.

a. State the appropriate null and alternate hypotheses.
$H_0 \leq 5 \hspace{10 pt} H_1 > 5$
b. Find the P-value.
```{r}
x = 6.5
s = 1.9
n = 8
t = (x - 5)/(s/sqrt(n))
print(t)
```
```{r}
t.values <- seq(-8,8,.1)
plot(x = t.values,y = dt(t.values, n-1), type = "l", lty = "dotted", ylim = c(0,.4), xlab = "t", ylab = "f(t)")
abline(v = t)
```

We are looking for the area to the right of the vertical line. That area indicates the probability of obtaining 6.5 or more given that $H_0$ is true.

```{r}
print(pt(t, n-1, lower.tail=FALSE))
```


c. Should the pump be put into service? Explain
Since the p-value is low, the $H_0$ hypothesis can be rejected. So the pump can be put into service.


4. A certain manufactured product is supposed to contain 23% potassium by weight. A sample of 10 specimens of this product had an average percentage of 23.2 with a standard deviation of 0.2. If the mean percentage is found to differ from 23, the manufacturing process will be recalibrated.

a. State the appropriate null and alternate hypotheses.
$$H_0 = 23 \hspace{10 pt} H_1  \neq 23$$

b. Compute the P-value.
```{r}
n = 10
x = 23.2
s = 0.2
t = (x - 23)/(s/sqrt(n))
print(t)
pvalue = pt(-t, df=n-1, lower.tail=TRUE) + pt(t, df=n-1, lower.tail=FALSE)
print(pvalue)
```
c. Should the process be recalibrated? Explain.


5. The article “Influence of Penetration Rate on Pen-etrometer Resistance” (G. Gagnon and J. Doubrough,Canadian Journal of Civil Engineering, 2011:741–750) describes a study in which twenty 2-L spec-imens of water were drawn from a public worksbuilding in Bridgewater, Nova Scotia. The mean lead concentration was 6.7 μg/L with a standard deviationof 3.9μg/L.

a. The Health Canada guideline states that the concentration should be less than 10μg/L. Can youconclude that the water in this system meets theguideline?
$$  H_0 \geq 10 \hspace{10pt} H_1 < 10 $$
```{r}
n = 20
x = 6.7
s = 3.9
t = (x - 10)/(s/sqrt(n))
print(t)
print(pt(t, df=n-1))
```
The water meets the guideline. The p-value is small

b. A stricter guideline is being considered, whichwould require the concentration to be less than7.5μg/L. Can you conclude that the water in thissystem meets this guideline?
$$  H_0 \geq 7.5 \hspace{10pt} H_1 < 7.5 $$
```{r}
n = 20
x = 6.7
s = 3.9
t = (x - 7.5)/(s/sqrt(n))
print(t)
print(pt(t, df=n-1))
```
The p-value is high so this systems does not meet the guildeline.


6. A new process for producing a type of novolac resinis supposed to have a mean cycle time of 3.5 hoursper batch. Six batches are produced, and their cycletimes, in hours, were
$$3.45, 3.47,3.57,3.52,3.40,3.63$$
Can you conclude that the mean cycle time is greater than 3.5 hours?
$$ H_0 \leq 3.5 \hspace{10pt} H_1 > 3.5 $$

```{r}
data = c(3.45, 3.47,3.57,3.52,3.40,3.63)
n = length(data)
x = mean(data)
s = sd(data)
t = (x - 3.5) /(s/sqrt(n))
print(t)
print(pt(t, df=n-1, lower.tail=FALSE))
```
There is not strong evidence to reject the null hypthoesis.

7. Specifications call for the wall thickness of two-literpolycarbonate bottles to average 4.0 mils. A qualitycontrol engineer samples 7 two-liter polycarbonate bottles from a large batch and measures the wall thick-ness (in mils) in each. The results are: 3.999, 4.037,4.116, 4.063, 3.969, 3.955, and 4.091. It is desired to test
$$ H_0: \mu=4.0 \text{ versus }  H_1:\mu \neq 4.0$$
a. Make a dotplot of the seven values.
```{r}
data = c(3.999, 4.037,4.116, 4.063, 3.969, 3.955, 4.091)
dotchart(data)
```
b. Should a Student’sttest be used to test H_0? If so,perform the test. If not, explain why not.
```{r}
data = c(3.999, 4.037,4.116, 4.063, 3.969, 3.955, 4.091)
qqnorm(data)
qqline(data)
```

The normal assumption is reasonable given the qqplot.

```{r}
data = c(3.999, 4.037,4.116, 4.063, 3.969, 3.955, 4.091)
n = length(data)
x = mean(data)
s = sd(data)
t = (x - 4)/(s/sqrt(n))
print(t)
print(pt(-t, df=n-1) + pt(t, df=n-1, lower.tail=FALSE))
```

There is not strong evidence to reject the null hyphotesis

c. Measurements are taken of the wall thicknesses of seven bottles of a different type. 
The measurements this time are: 
$$ 4.004, 4.225, 3.924, 4.052, 3.975,3.976,  4.041$$
Make a dotplot of these values.
d. Should a Student’s t test be used to test $H_0: \mu=4.0 \text{ versus } H_1:\mu \neq 4.0$? If so, perform the test. If not,explain why not.
```{r}
data = c(4.004, 4.225, 3.924, 4.052, 3.975,3.976, 4.0419)
dotchart(data)



qqnorm(data)
qqline(data)
```

There is an outlier, we should not use the student t-test

8. As part of the quality-control program for a catalyst manufacturing line, the raw materials (alumina anda binder) are tested for purity. The process requiresthat the purity of the alumina be greater than 85%.A random sample from a recent shipment of aluminayielded the following results (in percent):

$$93.2,  87.0,  92.1,  90.1,  87.3,  93.6$$

A hypothesis test will be done to determine whetheror not to accept the shipment.

a. State the appropriate null and alternate hypotheses.

$$ H_0: \mu \leq 85 \hspace{10pt} \mu > 85 $$

b. Compute the P-value.
```{r}
data = c(93.2,  87.0,  92.1,  90.1,  87.3,  93.6)
n = length(data)
x = mean(data)
s = sd(data)
t = (x - 85)/(s/sqrt(n))
print(t)
print(pt(t, df=n-1, lower.tail=FALSE))
```
c. Should the shipment be accepted? Explain.
Yes

9. The article “Approximate Methods for Estimating Hysteretic Energy Demand on Plan-Asymmetric Buildings” (M. Rathhore, A. Chowdhury, and S. Ghosh, Journal of Earthquake Engineering, 2011: 99–123) presents a method, based on a modal pushover analysis, of estimating the hysteretic energy demand placed on a structure by an earthquake.A sample of 18 measurements had a mean error of 457.8 kNm with a standard deviation of 317.7 kNm. An engineer claims that the method is unbiased, in other words, that the mean error is 0. Can you conclude that this claim is false?
```{r}
x = 457.8
n = 18
s = 317.7
t = x/(s/sqrt(n))
print(t)
print(pt(-t, df=n-1) + pt(t, df=n-1, lower.tail=FALSE))
```
